### 2024-05-21 paper/ms.tex
update ms

% upon AAS submission
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
% drafting / arxiv
\documentclass[11pt,twocolumn,tighten]{aastex63}
\turnoffedit

\usepackage{apjfonts}
\usepackage{url}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{amsmath,amstext,amssymb}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{xcolor, fontawesome}
\usepackage{color}
\usepackage{enumitem}

\newcommand{\red}{\color{red}}

\newcommand{\rprs}{{$R_p/R_{\star}$}}
\newcommand{\vsini}{{$V \sin i$}}
\newcommand{\kms}{{km\,s$^{-1}$}}
\newcommand{\gcc}{{g\,cm$^{-3}$}}
\newcommand{\rstar}{{$R_\star$}}
\newcommand{\rhostar}{{$\rho_\star$}}
\newcommand{\mearth}{{M$_\oplus$}}
\newcommand{\rearth}{{R$_\oplus$}}
\newcommand{\rsun}{{$R_\odot$}}
\newcommand{\msun}{{$M_\odot$}}
\newcommand{\bprp}{G_{\rm BP} - G_{\rm RP}}

\newcommand{\minus}{\scalebox{0.5}[1.0]{$-$}}

%%%%%%%%%%%%%%%%
% INSTITUTIONS %
%%%%%%%%%%%%%%%%
\newcommand{\caltech}{Department of Astronomy, MC 249-17, California Institute of Technology, Pasadena, CA 91125, USA}
\newcommand{\mitkavli}{MIT Kavli Institute and Department of Physics,
77 Massachusetts Avenue, Cambridge, MA 02139, USA}
\newcommand{\berkeley}{Astronomy Department, University of California,
Berkeley, CA 94720, USA}
\newcommand{\cmu}{Department of Statistics \& Data Science, Carnegie
Mellon University, Pittsburgh, PA 15213, USA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%
% VALUES %
%%%%%%%%%%
\input{vals_manuscript}

% number of stars monitored by Kepler (quoting Santos19/21)
\newcommand{\nkeplerstars}{$\approx$160{,}000}
% fraction of stars with rotation periods, with teff/logg in Berger+20
\newcommand{\fracstarswithprotwithbtwenty}{{$\approx$94\%}}
% complement of above
\newcommand{\fracstarswithprotwithoutbtwenty}{{$\approx$6\%}}
% fraction of KOIs that are not FPs with B+20 parameters
\newcommand{\frackoisnofpwithprotwithbtwenty}{{$\approx$92\%}}

%%%%%%%%%%%%%
% OFT-CITED %
%%%%%%%%%%%%%
\defcitealias{2013ApJ...775L..11M}{M13}
\defcitealias{McQuillan_2014}{M14}
\defcitealias{Mazeh_2015}{M15}
\defcitealias{Santos_2019}{S19}
\defcitealias{Santos_2021}{S21}
\defcitealias{Fulton_2018}{F18}
\defcitealias{Berger_2020a_catalog}{B20}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field}

\correspondingauthor{Luke G. Bouma}
\email{luke@astro.caltech.edu}

\received{---}
\revised{---}
\accepted{---}
\shorttitle{Kepler's Demographic Cliff}

\shortauthors{Bouma et al.}

\author[0000-0002-0514-5538]{Luke~G.~Bouma}
\altaffiliation{51 Pegasi b Fellow}
\affiliation{\caltech}
%\affiliation{51 Pegasi b Fellow}

\author{and coauthors, including but not limited to...}

\author{Lynne~A.~Hillenbrand}
\affiliation{\caltech}

\author[0000-0001-8638-0320]{Andrew W. Howard}
\affiliation{\caltech}

\author[0000-0002-0531-1073]{Howard Isaacson}
\affiliation{\berkeley}

\author[0000-0001-7967-1795]{Elsa~K.~Palumbo}
\affiliation{\cmu}

% <250 words
\begin{abstract}
  %
  Recent analyses of FGK stars in open clusters have helped clarify
  the precision with which a star's rotation rate and lithium content
  can be used as empirical indicators for its age.
  %
  Here we apply this knowledge to stars observed by Kepler.
  %
  Rotation periods are drawn from previous work; lithium is
  measured from new and archival Keck/HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  and \nplwgyroagenograzing\ planets for which our method is
  applicable.
  %
  We find that our rotation-based ages accurately recover ages of
  stars in open clusters; they also agree with
  $\gtrsim$90\% of the independent lithium ages.
  %
  The results include \nplyounggyrotwosigma\ planets
  younger than 1\,Gyr at 2$\sigma$ (\nplyounggyrothreesigma\ planets
  at 3$\sigma$).
  %
  More broadly, we find that the star formation rate in the Kepler field
  has dropped by a factor of \ratiosfr$\pm$\uncratiosfr\ over the past
  3\,Gyr.  This trend holds for both planet hosts and for the overall
  stellar sample.
  %
  This ``demographic cliff'' is a known feature of the Galaxy's star
  formation history, and it helps clarify the relative balance of
  young and old planets that have been detected to date using
  transits.
  %
\end{abstract}

\keywords{Stellar ages (1581), Planet hosting stars (1242), Field
stars (2103), Exoplanet evolution (491), Milky Way evolution (1052)}

\section{Introduction}
\label{sec:intro}

Exoplanet science is enjoying an age of abundance, fueled by the
discovery of thousands of worlds orbiting close to their host stars
\citep{Borucki10,2015JATIS...1a4003R}.  Most of these exoplanets are
billions of years old.  This fact leaves many gaps in our knowledge of
the exact physical and dynamical histories of these objects.  The
reason is that processes such as thermal cooling
\citep{2007ApJ...659.1661F}, atmospheric loss
\citep{2019AREPS..47...67O}, giant impacts
\citep{2014prpl.conf..595R}, and dynamical instabilities
\citep{2017MNRAS.470.1750I} are expected to be most efficient over
timescales of much less than 1\,Gyr.  For most known exoplanets, these
processes have run their course.

Young ($<$1\,Gyr) exoplanets therefore represent one means of building
the timeline for exoplanet evolution.  Informative individual
exemplars include V1298~Tau, an adolescent resonant chain of four
transiting puffy planets that is a likely precusor to the compact
multiplanet systems \citep{David_2019}, and HIP~67522, a Jupiter-sized
planet with a Neptunian mass (\citealt{Rizzuto_2020}; Thao et
al.~submitted).  Population-level analyses have similarly suggested
differences in the size distribution of young exoplanets relative to
their older counterparts
\citep{Berger_2020b_rpage,David_2021,Sandoval_2021,2023AJ....166..248C,2024arXiv240303261V}.

Discovering a young planet requires solving two problems: find the
planet, and measure the star's age.  Each problem admits a range of
solutions
\citep[e.g.][]{2008Sci...322.1348M,2012ApJ...756L..33Q,2024AJ....167..193T}.
In this article we will consider planets whose existence is previously
known from transits, and infer new stellar ages using rotation and
lithium.

To begin, imagine that the ages of nearby stars in the Galaxy are
uniformly distributed from 0--10\,Gyr
\citep[][]{2000MNRAS.318..658B,Nordstrom_2004}.  This approximation
clarifies the rarity of young stars: under this assumption,
$\approx$1\% of nearby stars have $t$$<$100\,Myr, and $\approx$10\%
have $t$$<$1\,Gyr.  Studies of currently forming protoplanets
\citep{2018A&A...617A..44K}, and of exoplanets evolving just after
disk dispersal \citep[e.g.][]{2022MNRAS.512.5067K}, are thereby capped
in their maximum achievable sample sizes by the tyranny of the
galactic star formation rate.

Despite this and other observational challenges, young close-in planet
discovery has matured over the past decade, primarily due to Kepler,
K2, TESS, and Gaia
\citep[e.g.][]{Meibom_2013,Mann_K2_25_2016,Curtis_2018,Livingston_2018,David_2019,Bouma_2020_toi837,Rizzuto_2020,Plavchan_2020,Newton_2021,Nardiello_2022,Barber_2022,Zhou_2022,Zakhozhay_2022,Wood_2023}.
The strategy pursued by most groups during the 2010s was to focus on
stars with known ages---obvious members of open clusters---and to
search them for transiting planets.  The resulting stellar (and
assumed planetary) ages are precise at the $\approx$10\% level.  A
more recent development, facilitated by the second Gaia data release,
is the idea that finding transiting planets is often easier than
finding the host stellar association \citep[e.g.][]{Tofflemire_2021}.

The ``cluster-first'' approach has a major limitation: most nearby
sub-gigayear stars are in the field.  Only $\lesssim$1\% of stars
within $\approx$500\,pc have been associated with their birth cluster
\citep[e.g.][]{Zari_2018,CantatGaudin_2020,Kounkel_2020,Kerr_2021}.
One can appreciate the problem by querying the NASA Exoplanet Archive
\citep[NEA;][]{2013PASP..125..989A} for transiting planets younger
than 1\,Gyr.  Imposing a precision requirement of say $t$$<$1\,Gyr at
2$\sigma$ precision gives $\approx$50 ``young'' transiting planets at
the time of writing.  Most of these young planets are in clusters, and
were found by Kepler, K2, or TESS.  However, these surveys have
discovered $\approx$5{,}000 planets in total.  Assuming a constant
star formation history, we would expect an order of magnitude more
sub-gigayear transiting planets.

This study aims to resolve two main questions.  First, how wrong is it
to assume a uniform age distribution for transiting planet host stars?
Second, where are the missing young planets?  We will find that
uniform is wrong by a factor of a few, and that stellar activity may
be a less significant bias than the challenge of precisely determining
ages for field stars.

Despite Kepler's main mission ending quite some time ago, the ages of
many Kepler planets remain uncertain.  While isochrone ages have been
calculated for Kepler stars \citep{Berger_2020b_rpage} and Kepler
Objects of Interest \citep[KOIs;][]{Petigura_2022}, isochrone ages are
most precise for stars whose luminosities and temperatures separate
them from the main sequence.  For $\lesssim$1\,Gyr stars, isochrone
ages are therefore limited.

Stellar rotation periods offer a promising alternative.  The idea of
using a star's spin-down as a clock has a rich history
\citep{Skumanich_1972,Noyes_1984,Kawaler_1989,Barnes03,Mamajek_2008,Angus_2015}.
Modern models yield ages precise to $\lesssim$30\% for FGK stars
between 1-4\,Gyr, and constraining age posteriors at earlier times
\citep{Bouma_2023}.  Physics-based models
\citep{Matt_2015,Gallet_Bouvier_2015,Spada_2020} are beginning to
connect these empirical relations to the evolution of stellar winds,
internal structure, and angular momentum transport.

In the context of Kepler, rotation-based ages have been reported for
planet-hosting stars since the early data releases
\citep{Walkowicz_2013,Reinhold_2015,Angus_2018}.  Recent studies
focused on rotation-based ages for Kepler have further explored
incorporating information from stellar kinematics
\citep{2021AJ....161..189L,2024AJ....167..159L}, and from light curve
features such as starspot amplitudes \citep{2023ApJ...952..131M}.  Our
analysis is motivated by a few factors that we believe constitute
improvements, particularly for the ages of sub-gigayear stars.   These
factors are as follows.
\begin{enumerate}[label={\it \roman*)},leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item The Kepler Object of Interest catalog, and our vetting of
    false positives within it, has reached maturity
    \citep[e.g.][]{Thompson_2018};
  %
  \item Measured rotation periods of FGK stars in open clusters now
    show not only the average evolution of $P_{\rm rot}(T_{\rm eff},t)$,
    but also how the astrophysical dispersion of stars around this average
    converges by the $\approx$700\,Myr age of Praesepe
    \citep[e.g.][]{Curtis_2019_ngc6811,Gillen_2020,Rampalli_2021,Fritzewski_2021,Rebull_2022,Dungee_2022,2023AJ....166...14B};
  %
  \item We can fit the open cluster data, and use these fits to infer
    ages by marginalizing over the range of ages that might explain
    any one star's rotation period \citep{Bouma_2023}.
    For Sun-like stars, this model is consistent with older
    calibrations \citep[e.g.][]{Mamajek_2008}, though with an improved treatment
    of uncertainty propagation.  For late G dwarfs and
    K dwarfs, the calibration differs by factors of a few;
  %
  \item We can identify binary stars at greater fidelity today than
    in the past, which can clarify otherwise problematic estimates of
    rotation-based ages.
  %
  \item New open clusters in various stages of dissolution have been
    found in the Kepler field
    \cite[e.g.][]{2019AJ....158..122K,Bouma_2022b,Barber_2022}.
    These clusters offer a sample of stars for which we can test the
    reliability of single-star age-dating methods.
\end{enumerate}

Independently, recent years have also yielded improvements in the
lithium age scale.  Lithium ages include two distinct regimes:
depletion boundary ages for M dwarfs and brown dwarfs in star
clusters, and the less precise decline-based ages for individual field
FGK stars \citep{Soderblom_2010}.  Here we focus on the latter
approach, which relies on the observation that the Li abundances of
partially-convective stars gradually decline as they age
\citep[e.g.][]{2005A&A...442..615S}.  The theoretical explanation for
this decline is debated
\citep[e.g.][]{1995ApJ...441..865C,2010ApJ...716.1269D,2019MNRAS.485.4052C}.
Empirical understanding has however improved due to the work by
\citet{Jeffries_2023}, who modeled the time evolution of the
\ion{Li}{1} 6708\,\AA\ equivalent width (EW) using a set of 6{,}200
stars in 52 open clusters.  Two-sided lithium ages are useful for
Kepler (FGK) stars between $\approx$0.03-0.5\,Gyr, with a strong
dependence on spectral type since the K dwarfs lose their surface
lithium much faster than G dwarfs.  The precision of the ages reported
by the \citet{Jeffries_2023} method in this regime are in the range of
0.3-0.5~dex.

We discuss our method for selecting the star and planet samples in
Section~\ref{sec:selection}, and describe the origin of our adopted
stellar parameters other than ages in Section~\ref{sec:stellarprops}.
We describe our age-dating methods in
Sections~\ref{sec:rotage} and~\ref{sec:liage}, and test them using
stars in clusters in the Kepler field.  The population-level
trends for both stars and planets are described in
Sections~\ref{sec:results} and~\ref{sec:disc}, and a few conclusions
are offered in Section~\ref{sec:conclusions}.

\section{Selecting the Stars and Planets}
\label{sec:selection}

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{st_params_M_G_vs_dr3_bp_rp.pdf}
		}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{st_params_adopted_logg_vs_adopted_Teff.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf The stars.}  Our analysis focuses on stars observed by Kepler
    with previously reported rotation periods (dark gray points).  The
    rotation periods were primarily measured by
    \citet{Santos_2019,Santos_2021}.  About half of these rotators are
    suitable for gyrochronology (black points), based on factors
    including their temperatures, surface gravities, non-binarity, and
    proximity to the main sequence (see Section~\ref{subsec:flags}).
    Some are ``confirmed'' or ``candidate'' Kepler Objects of Interest
    that meet an additional set of planetary quality criteria (blue
    points; Section~\ref{subsec:plflags}).  Surface gravities and
    effective temperatures were derived photometrically by
    \citet{Berger_2020a_catalog}.
	}
	\label{fig:stellarprops}
\end{figure*}

This work is focused on Kepler stars for which ages can be inferred
using either rotation, lithium, or both.  Such stars
are a minority of the \nkeplerstars\ Kepler targets.  Rotation periods
have been reported for roughly one in three Kepler targets
\citep[e.g.][]{McQuillan_2014,Santos_2021}.  High-resolution spectra
suitable for measurement of the \ion{Li}{1} 6708\,\AA\ doublet have
only been acquired for the Kepler objects of interest (KOIs), which
comprise a few percent of Kepler's targets.  In the following, we will
describe the set of stars that we adopt with measured rotation periods
(Section~\ref{subsec:rotsel}), the set of objects we adopt as hosting planets
(Section~\ref{subsec:planetsel}), and the subset of these with
high-resolution spectra suitable for lithium analysis
(Section~\ref{subsec:lithiumsel}).  Figure~\ref{fig:stellarprops}
provides a visualization of the various samples.

\subsection{Stellar Rotation Periods}
\label{subsec:rotsel}

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
		}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Rotation periods for Kepler target stars (left) and known
    planet hosts (right)}.  
    The gray lines are ``mean fits'' to the slow rotation sequences of
    open clusters.  The stellar sample (left) includes only apparently
    single stars near the main sequence with $\log g$$>$4.2,
    RUWE$<$1.4, and temperatures of 3800--6200\,K.  The color map has
    a linear stretch.  The planet-hosts (right) require the same
    stellar cuts, and include only the confirmed and candidate planets
    described in Section~\ref{subsec:plflags}.  An order of magnitude
    estimate for the number of young planets discovered by Kepler
    follows by counting planets below the mean rotational isochrones.
	}
	\label{fig:prot_vs_teff}
\end{figure*}

To select stars with rotation periods, we turn to previous work.  Many
investigators have derived rotation periods both specifically for KOIs
\citep{McQuillan_2013,Walkowicz_2013,Mazeh_2015,Angus_2018,David_2021},
and also for the broader set of all Kepler target stars
\citep{McQuillan_2014,Reinhold_2015,Santos_2019,Santos_2021}.  These
studies used a range of detection methods and selection functions.  We
are interested in understanding the age distribution of all Kepler
target stars irrespective of KOI status.  Out of these studies, the
only homogeneous analysis of both Kepler targets and KOIs were the
studies by \citet{Santos_2019,Santos_2021}, hereafter \citetalias{Santos_2019}
and \citetalias{Santos_2021}.  \citetalias{Santos_2019} and
\citetalias{Santos_2021} combined a wavelet analysis and
autocorrelation-based approach, and cumulatively reported rotation
periods for 55{,}232 main-sequence and subgiant FGKM stars.  They
included known KOIs and binaries in their analysis, and assigned them
specific quality flags, while removing transits and eclipses during
the stellar rotation measurement process.

%Studies of stellar rotation across the entire Kepler field include
%those by \citet{McQuillan_2014} (\citetalias{McQuillan_2014}),
%\citet{Reinhold_2015}, \citet{Santos_2019} (\citetalias{Santos_2019}),
%and \citet{Santos_2021} (\citetalias{Santos_2021}).
%\citetalias{McQuillan_2014} used an approach based on the
%autocorrelation function to detect 34{,}030 rotation periods for
%main-sequence Kepler targets cooler than 6500\,K, and excluded known
%eclipsing binaries and KOIs.  \citet{Reinhold_2015} used an iterative
%Lomb-Scargle based approach to analyze $\approx$40{,}000 stars with a
%variability range $R_{\rm var}>0.3$\% that were not known EBs, planet
%candidates, or pulsators; they reported primary rotation periods for
%24{,}124 of these stars.  \citet{Reinhold_2015} also reported
%secondary periods for two thirds of the stars with primary periods,
%which could be caused by differential rotation or finite spot
%lifetimes.  Finally, \citetalias{Santos_2019} and
%\citetalias{Santos_2021} combined a wavelet analysis and
%autocorrelation-based approach, and cumulatively reported rotation
%periods for 55{,}232 main-sequence and subgiant FGKM stars.
%\citetalias{Santos_2019} and \citetalias{Santos_2021} included known
%KOIs and binaries, and assigned them specific quality flags.  The
%rotation periods of KOIs have received considerable additional
%scrutiny
%\citep[e.g.][]{Walkowicz_2013,Mazeh_2015,Angus_2018,David_2021}.

We therefore adopt the results of \citetalias{Santos_2019} and
\citetalias{Santos_2021} as our default rotation periods.
\citetalias{Santos_2021} provided a comparison against
\citetalias{McQuillan_2014}; the brief summary is that the periods
agree for 99.0\% of the 31{,}038 period detections in common between
the two studies.  \citetalias{Santos_2021} classified the 2{,}992
remaining stars from \citetalias{McQuillan_2014} as not showing
rotation periods based on updated knowledge of contaminants
(e.g.~giant stars and eclipsing binaries) and visual inspection.  In
addition, \citetalias{Santos_2021} reported rotation periods for
24{,}182 main-sequence and subgiant FGKM stars that were not reported
as periodic by \citetalias{McQuillan_2014}.  Many of these reported
detections have lower variability amplitudes and longer periods than
those reported by \citetalias{McQuillan_2014}.

Analyzing the compilation of \citetalias{Santos_2019} and
\citetalias{Santos_2021} rotation periods for the KOI hosts, we
noticed that some known KOIs with rotation periods were missing.  This
is not surprising, in that the rotation periods of KOIs have received
considerably more scrutiny than those of ordinary Kepler stars.  We
therefore decided to split our subsequent analysis into a homogeneous
portion that used only the \citetalias{Santos_2019} and
\citetalias{Santos_2021} data, and an inhomogeneous portion that also
considered a broader set of available KOI rotation periods.  For the
latter portion, we first included 32 KOIs with orbital and rotation
periods within $\approx$20\% that had been
excluded %\footnote{Excluding stars in this category would impose a bias
%against young close-in planets, due to the natural commensurability
%between the rotation rates of young stars, and Kepler's increased
%sensitivity to short-period objects.  Although some of these KOIs are
%known false positives, our analysis excludes such objects at later
%steps.}%  As one motivating example, this subset of stars includes {\it
%e.g.}~Kepler~1643 (KIC~8653134), with $P_{\rm orb}$=5.34\,days and
%$P_{\rm rot}$=5.05\,days, which is known to be $\approx$40\,Myr old
%based on membership in RSG-5 \citep{Bouma_2022b}.  }
from the \citetalias{Santos_2019} and \citetalias{Santos_2021}
catalogs (A.~Santos, private communication).  We then incorporated an
additional \nnewdavidtwentyone\ rotation periods for KOI hosts that
\citet{David_2021} had described as either ``reliable'' or ``highly
reliable'' in their visual analysis of previously reported KOI
rotation periods from \citet{McQuillan_2013}, \citet{Walkowicz_2013}, \citet{Mazeh_2015}
and \citet{Angus_2018}.  Inclusion of these additional KOI rotation
periods is a supplementary measure aimed at completeness in our final
KOI age catalog; the provenances of the individual adopted periods are
noted in the relevant tables.

Finally, since our scope is focused on rotation-based ages, we
restricted our attention to stars with reported $P_{\rm
rot}$$<$45\,days.  The slowest-rotating FGK
stars in the open clusters used to
calibrate our gyrochronology model have $P_{\rm rot}$$\approx$35\,days.
Figure~\ref{fig:prot_vs_teff}
shows the resulting \nuniqstarsantosrot\ Kepler stars with rotation
periods compiled from \citetalias{Santos_2019},
\citetalias{Santos_2021}, and our extended KOI list.

To assess the statistical uncertainties of these rotation periods, we
compared our adopted \citetalias{Santos_2019} and
\citetalias{Santos_2021} periods with those reported by
\citet{McQuillan_2014}.  We found that for $P_{\rm
rot}\lesssim$15\,days, the two datasets agree at a precision of
$\lesssim$0.01$P_{\rm rot}$.  At longer periods of $P_{\rm
rot}\approx$30\,days, the agreement was typically at the
$\lesssim$0.03$P_{\rm rot}$ level, and the upper envelope of the
period difference increased roughly linearly with period.  Based on
this comparison, we adopted a simple prescription for the period
uncertainties, such that there are 1\% relative uncertainties below
$P_{\rm rot}=15$\,days, and a linear increase thereafter, with slope
set to require 3\% $P_{\rm rot}$ uncertainties at rotation periods of
30 days.

\subsection{Kepler Objects of Interest}
\label{subsec:planetsel}

We considered planets in the NEA cumulative KOI table as of 6 June
2023, which included the best knowledge available on any given planet
candidate while also incorporating human-based vetting.  These planets
represent a superset of those in the fully automated Q1-Q17 DR25 KOI
Table \citep{Thompson_2018}, which could be adopted in future work for
planet occurrence rate calculations.  This version of the cumulative
KOI table included \nkoisnofp\ objects that are either ``confirmed''
or ``candidate'' planets, after excluding known false positives.

\subsection{High Resolution Spectra}
\label{subsec:lithiumsel}

The final piece of our analysis involves assessing ages based on the
\ion{Li}{1} 6708\,\AA\ doublet.  We analyzed spectra
from the High Resolution Echelle Spectrometer (HIRES;
\citealt{vogt_hires_1994}) on the Keck I 10m telescope.  These spectra
were primarily collected through the California Kepler Survey
\citep{2017AJ....154..107P,2017AJ....154..108J,2017AJ....154..109F}.
We supplemented the existing archive with $\approx$10~hours of new
observations for 22 stars between Fall 2022 and Spring 2024.  These
stars were chosen to ensure that confirmed planets with rotational
evidence for ages below 1\,Gyr had spectra, since this is the age
range in which lithium is most likely to yield useful age constraints.

% stat from _estimate_new_lithium_timefrac.py
Lithium equivalent widths and abundances for the Kepler Objects of
Interest were already analyzed by \citet{2018ApJ...855..115B} for
roughly three quarters of the spectra in our sample.  However, new
spectra have since been acquired, and our approach and selection
function are different.  We therefore performed our own line width
measurements on
the reduced HIRES spectra.

We collected all blaze-corrected HIRES spectra from our group's
observations of non-false positive Kepler Objects of Interest with a
``multiple event statistic'' (MES, {\it koi\_max\_mult\_ev}) of at
least 10.  This yielded at least one spectrum for \nstwithspec\
stars hosting \nplwithspec\ planets.  About half of these stars have
measured rotation periods (\nstwithspecandprot\ stars and
\nplwithspecandprot, respectively).   For stars with multiple spectra
available, we analyzed only
the spectrum with the highest number of counts.  The resulting spectra
were acquired between 6 September 2009 and 16 May 2024, and
cumulatively comprise \nhireshours\,hours of open-shutter time.

We measured the lithium equivalent widths using a procedure adapted
from previous work \citep{Bouma_2021}.  Our stars of interest are FGK
stars, and so the continuum in the vicinity of the \ion{Li}{1}
6708\,\AA\ doublet is well-defined.  We Doppler-corrected the spectra
to a common reference wavelength by cross-correlating against a high
S/N template for Kepler-1698, chosen because $|\gamma|$$<$10\kms,
$T_{\rm eff}$$\approx$5000\,K and $v\sin i$$\approx$5\,\kms, which
puts it in the middle of our sample's temperature range and gives it
mild line-broadening.  We then trimmed the Doppler-corrected spectra
to a local window centered on the lithium line, using a window width
of 15\,\AA\ (we also considered 10\,\AA\ and 20\,\AA; the results were
consistent).  We continuum-normalized by fitting a third-degree
Chebyshev series, while excluding regions with absorption lines.  We
then numerically integrated the resulting spectrum using a
one-component Gaussian with free amplitude, width, and mean, and
estimated uncertainties on the line width through a Monte Carlo
procedure that bootstrapped against the local scatter in the continuum.

Our EW measurement approach does not correct for the neighboring
\ion{Fe}{1} 6707.44\,\AA\ blend.  To evaluate the accuracy and
precision of our measurement technique, after applying an initial
iteration on the \nlithiumstars\ KOIs with spectra, we compared our
lithium equivalent widths with those reported by
\citet{2018ApJ...855..115B}.  For the \nbergeroverlap\ stars in both
samples, we found broad agreement at $>$30\,m\AA, and significant
differences at $<$30\,m\AA\ because \citet{2018ApJ...855..115B}
required positive EWs, while we allow for statistically negative ones.  At
$>$30\,m\AA, there is however a small offset in our respective scales,
such that (B24-B18)$=$7.50$\pm$8.55\,m\AA.  This offset is caused by
our non-treatment of the iron blend.  We therefore directly subtract
this constant mean value (7.50\,m\AA) when calculating lithium-based
ages.

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{li_vs_teff_koi_X_JUMP_eagles.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Equivalent widths (EW) of the \ion{Li}{1} 6708\,\AA\ doublet
    for planet-hosting stars.} These measurements were made from
    Keck/HIRES spectra collected between 2009 and 2024.  Lines are the ``mean''
    isochrones from the same models as in Figure~\ref{fig:models}.  The
    intrinsic dispersion at $\gtrsim$1\,Gyr becomes much larger
    than changes in the mean.  Some stars with lithium detections do
    not have detected rotation periods, and vice-versa.
		\label{fig:li_vs_teff}
	}
\end{figure}

\section{Stellar Properties}
\label{sec:stellarprops}

% https://gea.esac.esa.int/archive/documentation/GDR3/Data_analysis/chap_cu8par/sec_cu8par_apsis/ssec_cu8par_apsis_gspphot.html

Our default source for stellar temperatures and surface gravities is
the Gaia-Kepler Stellar Properties Catalog (GKSP;
\citealt{Berger_2020a_catalog}).  The GKSP parameters were reported
for stars with ``AAA'' 2MASS photometry, measured parallaxes in Gaia
DR2,  and $g$-band photometry available from either SDSS or else Kepler-INT
\citep{2012AJ....144...24G}.  The parameters themselves were derived using
\texttt{isoclassify} \citep{2017ApJ...844..102H} to interpolate over
the MIST isochrone grids \citep{Choi_2016,2016ApJS..222....8D}, given
the SDSS $g$ and 2MASS $K_{\rm s}$ photometry, the Gaia DR2
parallaxes, and $V$-band extinction from the
\citet{2018MNRAS.478..651G} reddening map.  The resulting stellar
parameters are available for \fracstarswithprotwithbtwenty\ of the
\nuniqstarsantosrot\ Kepler stars with rotation periods.  For the
remaining \fracstarswithprotwithoutbtwenty\ of stars that lack
temperatures and surface gravities from
\citetalias{Berger_2020a_catalog}, we adopt the values reported by
\citet{Santos_2019} and \citet{Santos_2021}, which for these cases
primarily derive from the \citet{Mathur_2017} DR25 Kepler Stellar
Properties Catalog, and are mostly derived from photometry.  In the
planet sample, \frackoisnofpwithprotwithbtwenty\ of the
non-false-positive KOIs with rotation periods have parameters from
\citet{Berger_2020a_catalog}, and the remainder similarly are drawn
from DR25.

\citet{David_2021} compared the photometric
\citetalias{Berger_2020a_catalog} stellar parameters ($T_{\rm eff}$,
$R_\star$, [Fe/H]) against the spectroscopic parameters from
\citet{Fulton_2018}.  The temperature scales did show a few-percent
systematic difference, with \citetalias{Fulton_2018} quoting higher
temperatures than \citetalias{Berger_2020a_catalog} for mid K dwarfs,
and lower temperatures for early F dwarfs.  Our age analysis,
described below, assumes Gaussian uncertainties in effective
temperatures; we by default adopt the maximum of the two-sided
uncertainties reported by \citetalias{Berger_2020a_catalog} as our
symmetric temperature uncertainty.  Overall, the impact of systematic
uncertainties in the temperature scale on our age uncertainties is
generally smaller than that of the intrinsic scatter in the
calibration sequences of the open clusters.

\section{Ages From Rotation}
\label{sec:rotage}

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{gyromodeldispersion.pdf}
		}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{li_vs_teff_koi_X_S19S21dquality_eagles_showpoints_nodata.pdf}
		}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf The models.}
    Points represent $10^4$ draws from models that have been fitted to
    rotation periods \citep{Bouma_2023} and lithium equivalent widths
    \citep[EWs;][]{Jeffries_2023} of stars in open clusters.  Lines
    are the ``mean models'' at various ages.  The intrinsic dispersion
    in rotation and lithium about these mean models, which is what the
    models fit, sets the theoretical precision floor for the
    age-dating methods.    Additional sources of uncertainty,
    including measurement uncertainty, impose further limits on
    achievable precision.  These models are calibrated using open
    clusters younger than 4\,Gyr.  The displayed points assume a
    uniform distribution in temperature for visual clarity.  The sizes
    of the points are the same in each panel, so that low apparent
    density signifies greater dispersion around the mean.
		\label{fig:models}
	}
\end{figure*}

We calculate gyrochrone ages using \texttt{gyro-interp}
\citep{Bouma_2023}, which is a method designed to address the fact
that stars with the same mass and same rotation period can have very
different ages \citep[e.g.][]{Curtis_2019_ngc6811}.  The gyrochrone
age posterior should therefore incorporate this intrinsic
population-level scatter into the precision with which the age can be
inferred.  Figure~\ref{fig:models} highlights this problem,
particularly in regions where the 0.1-1\,Gyr stars overlap.

Using our effective temperatures and rotation periods, we ran
\texttt{gyro-interp} over an age grid of 0 to 4\,Gyr, linearly spaced
over 500 grid points.  \texttt{gyro-interp} itself is calibrated out
to 4\,Gyr, the age of M67
\citep[see][]{Dungee_2022,Gruner_2023}.  To estimate the
stellar spin-down rate as a function of time and stellar temperature,
the framework interpolates between reference open clusters using cubic
Hermite polynomials.  We calculate the probability of the
rotation-based age $t_{\rm gyro}$, given the observed rotation
periods, effective temperatures, and their uncertainties by
integrating Equation~1 of \citet{Bouma_2023} at the default grid
resolution.  This procedure marginalizes over the intrinsic age and
temperature dependent scatter that is observed in the cluster
sequences to yield statistical age uncertainties.

\subsection{Stellar Quality Flags}
\label{subsec:flags}
We calculated gyrochrone ages for all \nuniqstarsantosrotteffcut\
stars with reported rotation periods and effective temperatures of
3800--6200\,K.  To flag stars for which we suspect these ages may not
in fact be valid, we built a set of quality flags, which we condensed
into a single binary number, $Q_{\rm star}$.  When and how the flags
represented in this should be applied depends on the question being
asked.  If the goal is to construct a false-positive free sample, they
could all be applied.  If the goal is to construct a complete sample,
then consider the examples of Kepler~1627Ab ($\approx$40\,Myr) and
Kepler~51c ($\approx$625\,Myr).  The former has a high RUWE due to a
resolved binary companion \citep{Bouma_2022a}; the latter is on a
grazing orbit \citep{2014ApJ...783...53M}.  We leave selection for or
against such cases as a decision to the user.

Three assumptions must hold for a rotation-based age dating method
like \texttt{gyro-interp} to be valid.  {\it 1)} The evolutionary
state of the star must be well-specified by its temperature and its
age, {\it 2)} the star's spindown must not be influenced by binary
companions, and {\it 3)} the rotation period distribution for field
stars of a given temperature and age must be identical to that of
equally aged open clusters (metallicity differences, for instance, are
ignored).  A rephrasing of the first condition is that the star must
be ``near the main sequence'' because during the post main sequence
stellar temperatures change.  From $\approx$0.08 to $\approx$3\,Gyr,
the stars of interest in this work (3800--6200\,K,
0.5--1.2\,$M_\odot$) have temperatures constant to $\approx$1\%
\citep{Choi_2016}.  From $\approx$3--4\,Gyr, a 1.2\,$M_\odot$ star's
temperature drops from $\approx$6200\,K to $\approx$6000\,K
($\approx$3\%), because its outer layers expand as it begins hydrogen
shell burning as a subgiant.  We treat this issue in a manner
discussed in ``bit 9'' below.

{\it Temperature range (bit 0)}---We require stars to have effective
temperatures $T_{\rm eff}$ between 3800 and 6200\,K in order to report
gyrochrone ages \citep{Bouma_2023}.   Stars hotter than 6200\,K spin
down very slowly, if at all.  Stars cooler than 3800\,K do spin down
over gigayear timescales
\citep{2016ApJ...821...93N,2023ApJ...954L..50E,2024arXiv240312129C},
but the currently available open cluster data have yet to clarify
exactly when the intrinsic scatter in the population decreases.

{\it Surface gravity (bit 1)}---We flagged stars as possible subgiants
if they had $\log g < 4.2$.  We set this threshold by examining the
relevant flag from \citet{berger_2018_radii_evolnstates},
cross-matched against \citet{Berger_2020a_catalog}.

{\it Absolute luminosity (bit 2)}---We calculated the absolute Gaia
DR3 $G$-band luminosity, ignoring reddening, using the reported
apparent $G$-band magnitude and parallaxes.  We flag stars with
$M_{\rm G} < 3.9$ or $M_{\rm G} > 8.5$, corresponding to main-sequence
spectral types earlier than $\approx$F8V or later than $\approx$M0.5V
\citep{Pecaut_2013}.

{\it Known eclipsing binaries (bit 3)}---We flagged any stars reported
to be in the final Kepler eclipsing binary catalog (KEBC;
\citealt{2016AJ....151...68K}).

{\it Kepler-Gaia crossmatch quality (bits 4 and 5)}--- In situations
where we wanted to leverage Gaia DR3 data, we used M.~Bedell's 4$''$
Kepler-to-Gaia crossmatch %\footnote{\url{https://gaia-kepler.fun}, last
%accessed 15 April 2024.  We used 4$''$ rather than the 1$''$ match
%because 91 KOIs are omitted in the 1$''$ match, 23 of which have
%either ``candidate'' or ``confirmed'' status.  Almost all of these 23
%are M dwarfs with high proper motions.  Performing the same crossmatch
%at 4$''$, only two KOIs (KOI-3993 and KOI-6531), both known
%false positives, fail to yield Gaia DR3 crossmatches.}, 
%which is a
%crossmatch
of the NEA \texttt{q1\_q17\_dr25\_stellar} catalog with Gaia DR3 (accessible
online at \url{https://gaia-kepler.fun}).  The
separation distribution of the Kepler-Gaia DR3 crossmatches is such
that 99.2\% of candidate matches are within 1$''$.   We nonetheless
noted an upturn in the candidate match rate from 3-4$''$; such sources
are flagged using bit 4.  For KIC stars with multiple potential Gaia
matches within the 4$''$ radius, we adopted the brightest star as the
default match.  In most such cases this was unambiguous because there
is a large brightness difference between the primary and any apparent
neighbors.  However cases with multiple stars within 4$''$ within
$\Delta G$$<$$0.5$\,mag are noted using bit flag 5.

{\it Gaia DR3 non-single-stars (bit 6)}---The Gaia DR3
\texttt{non\_single\_star} column in the \texttt{gaia\_source} table
flags known eclipsing, astrometric, and spectroscopic binaries.  We
directly included this column.

{\it RUWE (bit 7)}---We inspected diagrams of the Gaia DR3
renormalized unit weight error (RUWE) as a function of other stellar
parameters, and flag stars with RUWE$>$1.4 as possible binaries.  Such
astrometric outliers can be either bona fide astrometric binaries, or
more often are marginally resolved point sources for which a
single-source PSF model provides a poor fit.

{\it Crowding (bit 8)}---We searched the Gaia DR3 point source catalog
for stars within 1 Kepler pixel (4$''$) of every target star.  While
such companions may not not be physically associated with the target
star, their presence can confuse rotation period measurements.  We
therefore flagged any stars with neighbors down to 10$\times$ the
brightness of the target star within this region ($\Delta G < 2.5$).

{\it Near the main sequence (bit 9)}---Many stars observed by Kepler
are far from the main sequence: this can be seen in both panels of
Figure~\ref{fig:stellarprops}.  This is a challenge for rotation-based
ages because {\it i)} point sources far from the main sequence include
unresolved photometric binaries, and {\it ii}) effects other than
magnetic braking influence the rotation periods of bona fide single
stars far from the main sequence.  The added concerns of
reddening and metallicity further complicate matters.  After exploring
various options, we settled on the orange locus in the $\log
g$--$T_{\rm eff}$ plane shown in Figure~\ref{fig:stellarprops}.  While
the exact details of how this locus is constructed are arbitrary\footnote{For purposes of reproducibility, 
	the method was as follows.
The orange line in Figure~\ref{fig:stellarprops} is	defined as
the maximum of two numerically-determined functions, $f_0(T_{\rm eff})$ and $f_1(T_{\rm eff})$.
We defined $f_0$ by fitting an
$N^{\rm th}$-order polynomial to the stars with rotation periods for which $t_{\rm
B20,iso}/(+\sigma_{t,{\rm B20,iso}})\approx 3$
and $T_{\rm eff}\in[3700,5900]$\,K, where
$t_{\rm B20,iso}$ was the isochronal age reported by
\citet{Berger_2020a_catalog}.  We let the order of the fit vary from
$N$=1 to 10, minimized the Bayesian Information Criterion,
and found $N$=6.  
By eye, this yielded a plausible locus for $T_{\rm eff}$$\lesssim$5300\,K.
However for F and G dwarfs, the resulting locus
allowed for stars that were `too evolved''; the density of KOIs 
was anomalously low in this region of the $\log g$ vs.\ $T$$_{\rm eff}$ plane.
We therefore visually selected KOIs that appeared
to be near the main sequence -- i.e.~ most of the blue
points in Figure~\ref{fig:stellarprops} -- and used them to fit a separate
polynomial $f_1$ through a similar BIC-minimization procedure,
and found $N$=3.
The portion of the orange locus from $\approx$5300--6200\,K
is set as $f_1 + c$, for $c$ a constant offset that we visually tuned to be
$\approx$0.1\,dex above the ``KOI main sequence''.
The exact break-point in temperature is automatically set by ${\rm max}(f_0, f_1+c)$.
We emphasize that for stars
surface gravities above the orange locus, we do report rotation based ages;
they are simply flagged as not being near the main sequence.  }
the general aim is to flag stars for which there is evidence based on
their location in the $\log g$--$T_{\rm eff}$ plane that our
gyrochronology model may not be reliable.  
%The rotation period
%distribution of sources excluded in this manner includes a relatively
%larger fraction of ``rapid rotators'' than the main sequence sources,
%likely due to a combination of astrophysical and systematic effects.
See for instance Figure~8a of \citet{2022AJ....164..137K} or
Figures~12 and 13 of \citet{2023ApJS..268....4F} for examples of how
the period distribution of this population differs from stars near the
main sequence.

{\it Candidate pulsators and close-in binaries (bit
10)}---\citeauthor{Santos_2021} included in a flag for candidate
``classical pulsators'' (e.g.\ Cepheids) and close-in binaries.
Visual inspection shows that although this flag selects many bona fide
objects in these classes, it also selects most $P_{\rm
rot}$$\lesssim$5\,day bona fide young planets.  These objects are
neither classical pulsators nor close binaries.  While we propagated
this flag into our set of quality flags, we therefore generally ignore
it.

{\it Not in the homogeneous stellar rotation sample (bit 11)}--- This
flag is set for the KOIs with rotation periods drawn from a source
other than \citet{Santos_2019,Santos_2021}.

{\it How do we use the bitmasks?} By default, we assume that a star
is suitable for gyrochronology if none of bits zero through nine
(inclusive) are raised.  For analyses that require a homogeneous
stellar rotation selection function, we also require bit 11 to be
zero.

\subsection{Planet Quality Flags}
\label{subsec:plflags}
Some planets are more reliably identified than others.  We use the
following additional quality indicators to assess the reliability and
utility of a planet, and assemble them into a separate bitmask,
$Q_{\rm planet}$.

% see Erik's 2022 discussion; might want to shift to include a cut on
% disposition score?
{\it Candidate reliability (bit 0)}---The NEA's Kepler Objects of
Interest Table includes both an overall planet-candidate disposition
status ({\it koi\_disposition}), as well as a disposition based only
on the Kepler data ({\it koi\_pdisposition}).  We require both to
include only ``planet candidates'' and ``confirmed planets''. 
%; we are generally less interested
%in the age distribution of planet candidates that are known false positives.

{\it Candidate S/N (bit 1)}---In any transit survey, the false
positive rate increases greatly toward the noise floor for planet
detection \citep[e.g.][]{2002ApJ...564..495J}.  We required a S/N in
excess of Kepler's usual 7.1$\sigma$ floor, through a cut on the
maximum ``multiple event statistic'' (MES, {\it koi\_max\_mult\_ev}):
we required ${\rm MES}$$>$10.

{\it Grazing planets (bit 2)}---Grazing objects, for which the impact
parameter $b$ is greater than $1-R_{\rm p}/R_\star$, often yield
biased planetary parameters \citep[e.g.][]{2022AJ....163..111G}.  For
large planets, they also include astrophysical false positives at
higher rates \citep{2016ApJ...822...86M}, in part due to the
size-impact parameter degeneracy.  We flag planet candidates as
potentially grazing if $b>0.8$, using the impact parameters reported
in the Markov Chain Monte Carlo (MCMC) fits from
\citet{Thompson_2018}.

\subsection{Sample Summary}
\label{subsec:tally}

The rotation periods for the star and planet samples are shown in
Figure~\ref{fig:prot_vs_teff}.  Out of the \nkeplerstars\ stars
observed by Kepler, requiring reported rotation periods and a reliable
DR3 crossmatch yields \nuniqstarsantosrot\ stars.  Further requiring
these stars to have temperatures of 3800--6200\,K, yields
\nuniqstarsantosrotteffcut\ stars.  Further requiring the stars to be
apparently single, near the main sequence, with $\log g$$>$4.2,
$M_{\rm G}>3.9$, and not flagged as eclipsing binaries yielded
\nuniqstarsantosallbutruwe\ rotators potentially amenable for
gyrochronology.  Finally, imposing RUWE$<$1.4 and requiring no
possible blends ($\Delta G$<2.5) within 4$''$ leaves
\nuniqstarsantosrotgyroappl\ stars for which gyrochronology is likely
to be valid.

In the planet sample, the cumulative KOI table included
\nnonfpcumkois\ confirmed and candidate planets, orbiting
\nnonfpcumkoihosts\ stars.  Assume that we apply all the same stellar
cuts, save for that on RUWE.  We can then require these confirmed and
candidate planets to have MES$>$10.  This combination of selection
criteria yield \nplwgyroagewithgrazingandhighruwe\ planets with
rotation-based ages orbiting
\nplhoststarwgyroagewithgrazingandhighruwe\ stars
(\nplhoststarwgyroagejusthighruwe\ of these stars with ``high'' RUWE).
If we additionally require the planets to be non-grazing ($b<0.8$), we
are left with \nplwgyroagenograzing\ planets orbiting
\nplhoststarwgyroagenograzing\ stars.

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.98\textwidth]{olympic_merged_lit_ages_vs_cluster_ages.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Literature age comparison.}
		Each point denotes a reported age,
		normalized by the age of a stellar ensemble in which the star
		is a candidate member.
		We used candidate cluster members as reported by previous studies based on spatial and kinematic
		information.  Although some field interlopers may be present in the 
		membership lists,
		any outliers can be compared between different methods on a
		relative basis.
		The number of stars in any given cluster follows by crossmatching every study against
		this fixed cluster list; certain methods report ages for more stars than others.
		Horizontal scatter is artificially added to visually
		clarify the statistical age uncertainties.
	}
	\label{fig:agescalecompone}
\end{figure*}

\section{Ages From Lithium}
\label{sec:liage}

Figure~\ref{fig:li_vs_teff} shows our measured equivalent widths for
the lithium 6708\AA\ absorption doublet, plotted over the mean
isochrone models from \texttt{EAGLES} \citep{Jeffries_2023}.  We
declare an upper-limit for plotting purposes if the 1$\sigma$ lower
limit on the equivalent width is below 10\,m\AA.  Our measured EWs
span -40 to 250\,m\AA; all negative EWs have uncertainties that are
statistically consistent with zero.

To calculate lithium ages, we use \texttt{EAGLES} (git commit
\texttt{ac09637}), which, similar to \texttt{gyro-interp}, is based on
a purely empirical interpolation approach.  \texttt{EAGLES} 
was calibrated on lithium measurements of stars observed by the
Gaia-ESO spectroscopic survey in 52 open clusters with ages spanning 2
to 6{,}000\,Myr \citep{Jeffries_2023}.  We adopt a linear prior in
age, spanning $10^6$ to $10^{10}$\,yr, and symmetrize our EW
uncertainties for purposes of interfacing with \texttt{EAGLES} by
taking the maximum of the measured positive and negative 1$\sigma$
uncertainty intervals.

In part because many of our EWs are upper limits, many of the lithium
ages are lower limits.  In such cases,  the inferred age posterior is
strongly influenced by the assumed intrinsic dispersion in
the lithium model, and by the prior.  In these instances, we
quote the 3$\sigma$ (99.7$^{\rm th}$ percentile) limits.

\section{Cluster Age Comparison}
\label{sec:litagecomparison}

Stars that formed in the same birth cluster 
are the gold standard for the astronomical age scale
\citep{Soderblom_2010}.
Before Gaia, the few $\gtrsim$1\,Gyr open clusters
known in the Kepler field had been visually identified by
\citet{1864RSPT..154....1H}.
Gaia proper motions and distances 
enabled the discovery of stellar ensembles that are typically more diffuse,
but which nonetheless
share a common age 
based on e.g.~isochrone, rotation, and lithium dating
\cite[e.g.][]{2019AJ....158..122K,Bouma_2022b,Barber_2022}.

In Figure~\ref{fig:agescalecompone}, we compare 
ages from the literature against the ages of stars in the richest
clusters in the Kepler field.
In particular, we drew ages for Kepler stars from \citet{Berger_2020a_catalog},
\citet{2021AJ....161..189L}, \citet{2023ApJ...952..131M}, and
\citet{2024AJ....167..159L}.
We followed any guidance available from these respective studies for
removing ages that were explicitly flagged as unreliable.

Cluster memberships are a delicate subject.
Figure~\ref{fig:agescalecompone} is showing reported ages for
a set of spatially and kinematically selected set of stars that could be cluster members, or
they could be field interlopers.  For NGC6811 and NGC6819, we
adopted candidate members from
\citet{2018A&A...618A..93C,CantatGaudin_2020} and
\citet{Kounkel_2020}. 
For Melange-3 we used the candidates reported by
\citet{Barber_2022} and required ``offset'' tangential
velocities below 2\,\kms.  For Cep-Her we
used candidate members from Table~2 of \citet{Bouma_2022b}, requiring
``weights'' exceeding 0.1.
Even with contaminants, we can compare
the relative age distributions derived by the different studies because in all
instances we are comparing reported ages against a fixed list of stars.

There are two main metrics for success in this test: accuracy
in whether the reported ages agree with the cluster age, and precision in
whether the reported age uncertainties agree with their
dispersion around the cluster age.

The point of Figure~\ref{fig:agescalecompone} is hopefully clear.
While previous studies have been able to derive accurate ages for
stars $\gtrsim$1\,Gyr old, the ages of sub-gigayear stars have historically been
overestimated by anywhere from 0.5--2\,dex.
The situation is particularly severe for the 40\,Myr stars in Cep-Her,
but it extends out to 150--300\,Myr stars in Melange-3 and Theia~520.
For isochrone and kinematic ages,
this is not surprising because these methods use stellar parameters that do
not appreciably change at $t$$\lesssim$1\,Gyr.
For the rotation-based ages quoted by
\citet{2023ApJ...952..131M} and \citet{2024AJ....167..159L}, the discrepancy
between their ages and the cluster age scale is likely caused by details of the calibration 
methodologies.
The physical spin-down model used by
\citet{2023ApJ...952..131M},
and the gyro-kinematic calibration ages used by \citet{2024AJ....167..159L},
may not reproduce
the rotation sequences of open clusters such as Praesepe, the Hyades, and the Pleiades.
The statistical age uncertainties yielded by these approaches
similarly may not marginalize over the intrinsic
scatter in rotation periods present in stars of a fixed mass at any given age.

%The rotation-based ages that we have calculated have a precision
%limited by the intrinsic scatter in the cluster rotation sequences.
%Within the age range accessible through the clusters, they seem to be
%mostly accurate,
%with the one exception being that our ages for stars in Cep-Her 
%have a typical median of $\approx$80\,Myr, rather than 40\,Myr.
%The reason is that 80\,Myr is the lower bound of 
%the \texttt{gyro-interp}  age grid; pre-main-sequence
%spin-up is outside of our calibration regime.

\section{Results}
\label{sec:results}

\begin{figure*}[!t]
  \begin{center}
    \leavevmode
    \subfloat{
    \includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240430_maxage4000.pdf}
    }

	\vspace{-0.35cm}
    \subfloat{
    \includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240430_maxage4000_preciseagesonly.pdf}
    }
  \end{center}
  \vspace{-0.6cm}
  \caption{{\bf Kepler's demographic cliff,} visible in the
  rotation-derived ages of stars (left) and planet hosts (middle) in
  the Kepler field.  The top row corresponds to all stars with
  temperatures of 3800--6200\,K for which we calculated rotation-based
  ages.  The darker lines impose quality cuts on factors such as
  binarity, crowding, and the star's evolutionary state.  The bottom
  row selects only stars with temperatures of 4400-5400\,K, which have
  relatively precise ages near 2\,Gyr due to their fast spin-down
  \citep[see][]{Bouma_2023}.  The statistical uncertainty for a mean
  individual star at 1, 2, and 3\,Gyr is shown, and is identical
  across each row.  Finally, the right panel compares the
  rotation-based age distributions against star formation histories
  derived using CMD fitting.  Completeness is near unity until
  $t$$\lesssim$3\,Gyr \citep{2022ApJ...937...94M}.
  \label{fig:hist_tgyro}
  }
\end{figure*}

\subsection{The Demographic Cliff}

The stellar ages and planetary properties of the systems discussed in
the previous sections are given in Tables~\ref{tab:planets}
and~\ref{tab:stars}.  Figure~\ref{fig:hist_tgyro} shows the gyrochone
age distributions of all the stars (left column) and the known
planet-hosts (middle column).  These ``histograms'' are sampled from
the posterior probability distributions by drawing ten random samples
from each posterior, and then computing the re-weighted
histogram of the resulting samples.  The uncertainties assume Poisson
statistics.  We truncated the plots near $4$\,Gyr since this is the
cluster age out to which our interpolation-based method is calibrated.
We refer to the overall slope as a ``demographic cliff'' in reference
to an impending human phenomenon forecasted to happen in coming years.

The age distributions of Kepler stars and planet hosts are
nearly indistinguishable.  A simple diagnostic of whether the two
distributions are drawn from the same underlying distribution is the
Kolmogorov-Smirnov test;  we calculated this statistic in a manner
that accounts for the Poisson uncertainties by performing 100 random
draws of 70\% of the stars from each sample.  The resulting median
$p$-values were $\approx$0.01 for both rows of
Figure~\ref{fig:hist_tgyro}.  This degree of similarity suggests that
the ``common wisdom'' that young transiting planets are hard to detect
because their host stars are photometrically variable requires some
nuance.  If true, this statement only holds for the very youngest
stars, since at the age resolution of this study, there appears to be
little to no selection bias against sub-gigayear planets.

An important separate selection effect concerns Kepler's ability to
detect the rotation signals.  \citet{2022ApJ...937...94M} studied this
issue, and found for Sun-like stars that the fraction of rotation
signals detectable by Kepler is near unity up to $\approx$3\,Gyr, and
that it drops to almost zero by $\approx$5\,Gyr .  This trend is
opposite in functional form to our derived age distributions.

We can quantify the relative counts of stars as a function of age by
labelling stars between 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr as ``young'',
``intermediate-age'', and ``old''.  A simple counting exercise from
Figure~\ref{fig:hist_tgyro} tells us that there are \ratioobtoybstars\
times as many old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet
hosts as young planet hosts.  Focusing only on the tails of the
distributions (0-0.3\,Gyr and 2.7-3\,Gyr), the implication is that the
formation rate of stars in the Kepler field decreased by a factor of
\ratiosfr$\pm$\uncratiosfr\ over the past three billion years.

In terms of detected planet counts, our rotation-based ages for the
Kepler sample yield \{\nplyounggyro, \nplmidgyro, \nploldgyro\}
detected planets in the 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr bins.

\subsection{Planets Younger Than One Billion Years}

Sub-gigayear planets hold particular weight in studies of planet
evolution.
While our catalog has \nplyounggyro\ confirmed planets
with median ages below 1\,Gyr, requiring $t_{\rm gyro}$$<$1\,Gyr at
2$\sigma$ confidence yields \nplyounggyrotwosigma\ planets orbiting
\nplhostsyounggyrotwosigma\ stars.  
The most notable of these planets can be split into two classes: those
in clusters, and those with independent and consistent rotation and
lithium-based ages.

\subsubsection{Planets in Clusters}
\label{subsec:clusterplanets}

Four stellar ensembles in the Kepler field have to date yielded a
total of fourteen transiting planets.  Our analysis blindly recovers
the youth of these planets.

{\it Cep-Her}---The $\approx$40\,Myr Cep-Her complex currently has its
most constraining age measurement from relative isochrone dating of
pre-main-sequence M dwarfs; it contains Kepler-1627Ab, Kepler-1643b,
Kepler-1974b and Kepler-1975\,Ab \citep{Bouma_2022a,Bouma_2022b}.
Different sub-groups of the complex may vary in age by $\approx$50\%
(\citealt{Bouma_2022b}; Kerr~et al.,~submitted).  The planets
themselves are all on close-in orbits (5-25\,days), with sizes from
2-4\,$R_\oplus$.  The rotation, lithium, and cluster ages all agree
for Kepler-1627A, Kepler-1974, and Kepler-1975A.  For Kepler-1643, the
rotation and cluster ages agree, and the lithium age is 2.0$\sigma$
above the cluster age reported by \citet{Bouma_2022b} for the RSG-5
subgroup.

{\it MELANGE-3}---\citet{Barber_2022} reported a
$105\pm10$\,Myr association in the Kepler field and derived its age
using rotation, lithium, and the isochronal pre-main-sequence.
Kepler-970 and Kepler-1928 harbor two transiting mini-Neptunes inside
this association.  Our rotation-based
age-dating independently recovers these systems, with $t_{\rm
gyro}=176_{-40}^{+120}$\,Myr and $144_{-88}^{+104}$\,Myr for
Kepler-970 and Kepler-1928 respectively.  The former rotation-based
age being slightly older than the suggested Pleiades age for the
assocation agrees with Figure~4 of \citet{Barber_2022}.

{\it Theia~520}---We derive rotation-based ages of 330-350\,Myr for
Kepler-968 and Kepler-52.  These stars were reported by
\citet{2019AJ....158..122K} to be members of an association called
Theia~520 based on their spatial and kinematic correlations with
nearby stars.  The core of this structure is also known as UBC-1
\citep{2018A&A...618A..59C}; since both Kepler planet hosts are in
diffuse regions of the cluster, we use the name from
\citeauthor{2019AJ....158..122K}.  Based on isochrone-fitting,
Theia~520 is $247^{+64}_{-51}$\,Myr old \citep{2019AJ....158..122K}.
An independent analysis by \citet{2024A&A...681A..13F} using g-mode
asteroseismology, gyrochronology, and isochrone fitting found
consistent results.  Ongoing work by J.~Curtis et al.~broadly finds
that the two planet-hosting stars are indeed members of the diffuse
population.  Given the cluster-level and individual-star evidence,
this makes Kepler-968 and Kepler-52 the youngest multiplanet systems
currently known from the primary Kepler mission.

{\it NGC~6811}---\citet{Meibom_2013} reported the discovery of
Kepler-66 and -67b, two mini-Neptunes in NGC~6811 ($\approx$1\,Gyr).
We report rotation-based ages for these systems of
\kepsixsixtgyro\,Myr and \kepsixseventgyro\,Myr for Kepler-66 and
-67b, respectively.  Kepler-66 ($T_{\rm eff}$$\approx$5900\,K, $P_{\rm
rot}$$\approx$10.4\,days) has a lower precision and a more asymmetric
posterior because the spindown rates of stars this hot slow down after
the first gigayear.  Kepler-67 ($T_{\rm eff}$$\approx$5100\,K, $P_{\rm
rot}$$\approx$10.3\,days) is marginally hotter than the later K dwarfs
that have ``stalled'' spindown from 0.7 to 1\,Gyr, enabling its
rotation period to be diagnostic of the stellar age.

\begin{figure}[!t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_anyyoung_anyyoung.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf Sizes, orbital periods, and rotation-based ages of transiting
    exoplanets younger than one billion years.} Stars denote
    \nplyounggyrotwosigmanograzingnoruwe\ planets that are newly
    characterized in this work to have $t_{\rm gyro}$<1\,Gyr at
    $2$$\sigma$; circles are planets from the literature meeting the
    same age requirement, but with a heterogeneous set of age
    provenances.  Gray points are transiting planets from the NASA
    Exoplanet Archive older than 1\,Gyr.  If one were to instead
    select for precise ages ($t/\sigma_{\rm t}>3$), this would hide
    $\approx$ten $<$0.5\,Gyr planets, and add $\approx$fifteen
    0.5-1\,Gyr planets.  For this plot, we require our Kepler stellar
    hosts to have no quality flags raised, and for the planets to
    similarly no have quality flags raised. 
    \label{fig:rp_period_age_results}
  }
\end{figure}

\subsubsection{Comparing Rotation and Lithium Ages}

We compared the independent rotation and lithium-based ages of the
known planet hosts using the point-estimates of the available
posteriors.  There are three cases of interest:

\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item Two-sided posteriors for both $t_{\rm gyro}$ and $t_{\rm Li}$:
    if their values were consistent within 2$\sigma$, we judged them
    to be consistent; if they were consistent within 2-3$\sigma$, we
    suggest them to be ``maybe'' consistent.  Otherwise, they
    disagree.
    %
  \item If $t_{\rm Li}$ is a lower limit, we compare this lower limit
    with the $1\sigma$ upper limit from $t_{\rm gyro}$; if the two
    overlapped, we judge the age estimates to be consistent.
    %
  \item If $t_{\rm Li}$ provides a two-sided posterior, and no
    rotation is reported, we judge the age estimates to be
    inconsistent.  This is because this lithium constraint can only be
    provided for G and K stars $\lesssim$2\,Gyr old, and Kepler should
    have been sensitive to the rotation periods of such stars.
\end{enumerate}

In the entire planet sample (i.e.~without imposing any quality cuts) this
yielded \allagesyesconsistent\ consistent cases,
\allagesmaybeconsistent\ ``maybe consistent'' cases, and
\allagesnoconsistent\ inconsistent cases.  Rephrased, in the overall
sample, \fracconsistentallages\% of stars have consistent rotation and
lithium-based ages (\fracpotentiallyconsistentallages\% potentially
consistent).

Appendix~\ref{app:inconsistent} describes all cases of ``discrepant''
confirmed planets for which a sub-gigayear age is reported by at least
one age indicator.  All but two systems are either evolved stars or
else unresolved binaries, and are automatically flagged as such.  The
two exceptions are Kepler-786 (KOI-1199), and Kepler-1 (TrES-2).
Kepler-786 is an early K dwarf with $t_{\rm
Li}$=\kepseveneightsix\,Myr, which would imply that rotation should be
detectable.  However no photometric rotation signal is detected; other
age indicators in the spectrum similarly suggest that the star may not
in fact be young.  TrES-2 is therefore a ``lithium-young'' system for
which we believe the high lithium EW may be a statistical fluke.

\subsubsection{New Young Planets}

Let us restrict our attention to blemish-free stars, with blemish-free
``confirmed'' planets: $Q_{\rm star}$ must not have any of bits zero
through nine raised, and $Q_{\rm planet}$ must have no quality flags
raised at all.  Under this constraint, our results include
\ltonegyrhighqconfirmedtwosided\ systems younger than 1\,Gyr with
two-sided $t_{\rm gyro}$ and $t_{\rm Li}$, and
\ltonegyrhighqconfirmedonesided\ cases for which $t_{\rm gyro}$
provides an age constraint while $t_{\rm Li}$ is one-sided.

These planets are shown in Figure~\ref{fig:rp_period_age_results}.
They include \nearthshighqconfirmed\ Earth-sized or smaller planets,
no Jupiter-sized or larger planets, \nsubsaturnshighqconfirmed\
``super-Neptune'' ($>$4\,$R_\oplus$) sized planets,
\nminineptuneshighqconfirmed\ mini-Neptunes above the radius valley
(as parametrized by \citealt{2018MNRAS.479.4786V}), and
\nsuperearthshighqconfirmed\ super-Earth sized planets below the
radius valley.  While most of these planets are on orbital periods
below 50\,days, \nlongperiodhighqconfirmed\ are on more distant
orbits.

A few new highlights from the high quality stars and planets include
Kepler-1529, Kepler-1565, Kepler-1312, and Kepler-1629.  Kepler-1529b
is a $\approx$100$\pm$50\,Myr mini-Neptune with well-measured rotation
and lithium ages.  Kepler-1565b is an analogous $\approx$170-230\,Myr
super-Earth.  Kepler-1312 is a $t_{\rm gyro}$=\kepthirteentwelve\,Myr
near-Solar analog with two transiting planets: an Earth-sized planet
on a one-day orbit, and a mini-Neptune on a five-day orbit.  The
Earth-sized planet, Kepler-1312c, along with the near-identical
Kepler-1561 b ($t_{\rm gyro}$=\kepfifteensixone\,Myr), rank among the
youngest Earth-sized planets currently known, comparable to systems
such as HD~63433d \citep[1.1\,$R_\oplus$,
$414$$\pm$$23$\,Myr;][]{2024AJ....167...54C} and TOI-1807~b
\citep[1.3\,$R_\oplus$, 180$\pm$40\,Myr;][]{2021AJ....162...54H}.
Kepler-1629~b, with a size just two-thirds that of Earth
(0.67$\pm$0.05\,$R_\oplus$), is marginally older, also based on
rotation (\kepsixteentwonine\,Myr).  The lithium age from a
reconnaissance TRES spectrum confirms the point.

%  Kepler-1529b, a 5.3\,day 2.0\,$R_\oplus$ mini-Neptune is $\approx$100\,Myr old.
%  Kepler-1930b, a 13.0\,day 1.5\,$R_\oplus$ super-Earth on a grazing orbit is $\approx$120-180\,Myr old.
%  Kepler-1313b, doubly unlucky, a 3.8\,day 1.7\,$R_\oplus$ super-Earth on a grazing orbit is $\approx$185\,Myr old.
%  Kepler-1521b, a 47.2\,day 2.3\,$R_\oplus$ mini-Neptune is 180-210\,Myr old.
%  Kepler-1467b, a 47.1\,day 3.1\,$R_\oplus$ mini-Neptune is 230-240\,Myr old.
%  Kepler-1565b, a 1.5\,day 1.3\,$R_\oplus$ super-Earth is 180-230\,Myr old.
%  Kepler-971b, a 9.6\,day 3.9\,$R_\oplus$ Neptune-sized planet is 300-370\,Myr old.
%  
%  Kepler-1933b is a 4.9\,day 1.0\,$R_\oplus$ Earth-sized planet.  Its
%  rotation-based age of $304^{+96}_{-112}$\,Myr agrees within $1\sigma$ of
%  its lithium age, though at half the relative uncertainty.  Based on
%  these lines of evidence, this system may be the youngest Earth-sized
%  planet known, comparable to systems such as HD~63433d
%  \citep[1.1\,$R_\oplus$, $414\pm23$\,Myr][]{2024AJ....167...54C}.
%  
%  Kepler-1532b is a 1.1\,day 1.3\,$R_\oplus$ super-Earth.  Rotation points to a $\approx$400\,Myr age, and lithium is consistent.
%  Kepler-955b is a 14.5\,day 2.7\,$R_\oplus$ mini-Neptune.  Rotation points to a $\approx$400\,Myr age, and lithium is consistent.
%  Kepler-1481b, a 5.9\,day 1.1\,$R_\oplus$ super-Earth has a
%  rotation-based age of $\approx$400\,Myr, and a $2\sigma$ lithium
%  detection that, upon visual inspection of the spectrum, is
%  believeable.  The implied lithium age is marginally consistent with
%  the rotation-based age, although perhaps with some tension.
%  Kepler-447b is supposedly a 32\,$R_\oplus$ planet (grazing) on a
%  7.8\,day grazing orbit.  Rotation and lithium independently find a
%  420-440\,Myr age.
%  Kepler-1561b is a 0.9\,$R_\oplus$ sub-Earth, on a 1.0\,day orbit.
%  Rotation points to a 420\,Myr age.  The detected EW$_{\rm Li}$ is
%  consistent, although given the star's temperature, not particularly
%  constraining.
%  The same is true for the 1.9\,$R_\oplus$ and 5.8\,day super-Earth Kepler-523b, and the 
%  1.0\,$R_\oplus$, 12.2\,day Earth-sized Kepler-1482b.

\section{Discussion}
\label{sec:disc}

\subsection{The Thin Disk's Star Formation History}

Most of the stars observed by Kepler are in the thin disk.  This can
be verified by following \citet{Gaia_2018}, and denoting stars with 2D
tangential velocities $v_{\rm T}$$<$40\,\kms\ as thin disk members,
and those with 60$<$$v_{\rm T}$$<$150\,\kms as thick disk members.
For all stars in the KIC, this yields 95{,}694 thin and 54{,}039 thick
disk stars respectively.  The thin disk stars have a threefold larger
detected rotation period fraction: 35{,}675 rotators are in the thin
disk, while 7{,}312 are in the thick disk.  Further imposing the
gyrochronology quality flags discussed in Section~\ref{subsec:flags}
yields 17{,}755 thin-disk stars for which gyrochronology is nominally
applicable, and 2{,}462 thick-disk stars.

Classifying Kepler stars in this manner helps connect them to the
broader literature focused on the history of star formation in the
Galaxy.  On extragalactic scales, the star formation rate in spiral
galaxies with mass comparable to the Milky Way peaked $\approx$10\,Gyr
ago, and has since decreased by an order of magnitude
\citep[e.g.][]{2004Natur.428..625H,2006ApJ...651..142H}.  For our own
galaxy, many investigators have worked to measure similar star
formation histories, through both CMD fitting of resolved stellar
populations
\citep[][]{2019A&A...624L...1M,2020NatAs...4..965R,2021MNRAS.501..302A,2022Natur.603..599X}
and by modeling the local white dwarf luminosity function
\citep[e.g.][]{2019ApJ...878L..11I}.  \citet{2019A&A...624L...1M} and
\citet{2019ApJ...878L..11I} for instance focused on stars (and white
dwarfs) within a few hundred parsecs, and both found a peak in the
local star formation rate 2-3\,Gyr ago.  \citet{2020NatAs...4..965R}
focused on stars in a wider 2\,kpc bubble, and reported three local
maxima in the thin disk's star formation history 5.7, 1.9, and
1.0\,Gyr ago, which they associate with pericenter passages of the
Sagittarius satellite galaxy.

The right column of Figure~\ref{fig:hist_tgyro} compares our derived
age distribution against star formation histories (SFHs) reported by
\citet{2019A&A...624L...1M} and \citet{2020NatAs...4..965R} based on
CMD fitting.  The overall slope of both SFHs broadly agrees with what
we find from rotation-based ages.  The \citet{2019A&A...624L...1M}
data in particular seem entirely consistent, after accounting for the
statistical uncertainties reported in that study.  Regarding the
episodic star formation bursts reported by
\citet{2020NatAs...4..965R}, they are not apparent in our overall FGK
star sample (top row).  However, rotation-based ages are most precise
for $\approx$G8V-K4V dwarfs older than $\approx$1.5\,Gyr
\citep{Bouma_2023}.  The lower row of Figure~\ref{fig:hist_tgyro}
selects these stars with a temperature cut, and may show a hint of the
$\approx$2\,Gyr spike reported by \citet{2020NatAs...4..965R}.  The
$\approx$1\,Gyr spike however is not recovered.

% compare against Berger+20, 
% note their sentence
% "Encouragingly, the distribution also qualita- tively matches the red giant asteroseismology-derived age distributions in Silva Aguirre et al. (2018) and Pinsonneault et al. (2018), as well as the rotation-based ages in Claytor et al. (2020) and the Galactic Archaeology with HERMES–Gaia ages in Buder et al. (2019)."
%
% also compare against Mor2019 and their literature assessment
% "Our findings that the thin disc SFH does not follow a sim- ple decreasing shape until the present are in good agreement with Snaith et al. (2015), and Haywood et al. (2016, 2018) who found, using data with metallicities and assuming a fixed IMF, the existence of an SFR quenching followed by a reactivation. Kroupa (2002a), using stellar kinematics, found the SFH to behave similarly. The relative maximum of the SFR that we find at 2–3 Gyr ago is compatible with the results of Vergely et al. (2002) and Cignoni et al. (2006) that, using Hipparcos data in a sphere of 80 pc around the Sun and assuming a fixed IMF, found maximum peaks at 1.75–2 Gyr ago and 2–3 Gyr ago, respec- tively. Recently, Bernard (2018), in a tentative work using TGAS data, pointed towards the existence of a relative maximum also located 2–3 Gyr ago."

%\subsection{Isochrone Age Comparison}

%\subsection{Lithium Age Comparison}

\subsection{Caveats \& Limitations}

Our most constraining ages tend to come from one source of
information: rotation.  Factors other than stellar age and mass can
influence stellar rotation rates.  Binarity is one example: even
intermediate-separation ($\sim$10\,AU) binaries are biased toward
rapid rotation \citep[e.g.][and many studies thereafter]{Meibom_2007}.
Metallicity may also be relevant
\citep{2020MNRAS.499.3481A,2024arXiv240500779S}.  An additional caveat
is that our rotation-based ages used photometric effective
temperatures (Section~\ref{sec:stellarprops}), even though
spectroscopic temperatures are available for the planet-hosting
subset.  This decision was driven by a desire for homogeneity 
irrespective of planet-hosting status.  It may however have
incurred small systematic errors in the implied ages.

We were interested in not only the overall age
distribution of the Kepler field, but also in the reliability of
individual ages for young stars known to host planets.  
We assessed this reliability using
quality
bitmasks (Table~\ref{tab:planets}), 
a comparison against open cluster ages (Section~\ref{sec:litagecomparison}),
and a comparison between rotation and lithium ages.
The open cluster comparison suggested that our ages were generally
accurate at $\ll$0.5\,dex across 0.04--2.5\,Gyr, comparable to their
stated precision level.
Nonetheless,
outliers do exist in the upper-left panel of Figure~\ref{fig:agescalecompone},
typically at most 10\% of the population in each cluster.
These stars could be either field star contaminants, or else anomalous
stars whose physical rotation histories were altered by processes not
captured by our statistical uncertainties.
Similarly, although
we found $t_{\rm gyro}$ and $t_{\rm Li}$ to agree
for 
\ltonegyrhighqconfirmedtwosided\ planets orbiting ``high-quality''
stars, we did find
one system, Kepler-786, with a precise lithium age
(\kepseveneightsix\,Myr), but no rotation signal.

Other age indicators may help in verifying the ages of the
$\lesssim$3\,Gyr stars that are the focus of this work.  Specifically,
chromospheric emission in the X-ray, \ion{Ca}{2} HK line, and the UV
can serve as an age tracer
\citep{Mamajek_2008,2014MNRAS.441.2361V,2024ApJ...960...62E}.
However, these age indicators are all in a sense ``rotation-powered''.
The dynamo converts kinetic energy into magnetic energy, which is
emitted through these chromospheric pathways.  We did not attempt to
incorporate these indicators into this study due to concerns regarding
sensitivity, homogeneity, and the question of whether they in fact
provide age information that is truly independent from rotation.

A broader question, beyond the scope of this work, concerns the ages
of stars $\gtrsim$4\,Gyr old.  For
age-dating models based only on rotation rates, or else on a
combination of rotation and spin-modulation amplitude
\citep{2023ApJ...952..131M}, an important challenge is that the
detectability of the rotation signals for 3-10\,Gyr stars, especially
Sun-like stars, is at the limits of the current data
\citep{2022ApJ...937...94M}.  In the ``old star'' regime, methods
using either asteroseismology
\citep{vanSaders_2016,2024ApJ...962..138S}, or else using both the
evolution of stellar luminosity and stellar rotation
\citep{Angus_2019} seem the most capable of providing useful age
constraints for single stars.  Incorporating kinematics
\citep{2021AJ....161..189L} seems likely to remain most useful for
stellar ensembles, due to the stochasticity of kinematic heating.

\subsection{Future Directions}

{\it Occurrence Rates}:
The age-dependent trends predicted for exoplanet populations, such as
the Kelvin-Helmholtz cooling of mini-Neptunes \citep{Gupta_2019},
time-dependent carving of the photoevaporation desert
\citep{Owen_Lai_2018}, and the time evolution of the radius valley
\citep{Rogers_2021} can be explored using our data.  We defer this
analysis to a separate publication.
Some care is required since beyond age,
exoplanet demographics also depend on stellar metallicity and mass
\citep[e.g.][]{Petigura_2018}.

{\it Field star ages from other surveys}:
While this study focused on
Kepler, other photometric surveys (e.g.~K2, TESS, HAT, WASP, NGTS, ZTF,
ATLAS) open opportunities for age-dating a far broader set
of stars and planets.
Future prospects also include PLATO
\citep{Rauer14}, Earth 2.0 \citep{2022arXiv220606693G}, and the Roman
Galactic Plane Time Domain Survey.
The long-baseline surveys 
that specifically target different regions of the Galaxy
seem most likely to yield new insight into its star
formation history.

{\it Searches for new associations}:
Some of the youngest systems in Table~\ref{tab:planets} have been
linked to their birth clusters, while others have not. A systematic
search for the birth clusters of the youngest ``field stars'' could
combine spatial and kinematic information from Gaia with rotation
measurements from TESS. Given the $\sim$100\,Myr decoherence time of
young clusters, the youngest "field stars" may have easily
identifiable young neighbors.

\section{Conclusions}
\label{sec:conclusions}

We began this article with two questions: how wrong is the assumption
of a uniform age distribution for stars in the galactic thin disk?
And why are only $\approx$50 sub-gigayear transiting planets known,
rather than the $\approx$500 that would be expected under the
assumption of a uniform star formation history for the
$\approx$5{,}000 known planets?

Our approach to answering these questions was to curate a sample of
stellar rotation periods, lithium equivalent widths, and temperatures
using archival and new data.  We derived new ages using empirical
methods that have been calibrated against an extensive set of open
cluster data.

Our analysis blindly recovered all 14 known Kepler planets in
clusters, with ages in agreement with the more precise cluster-based
ages.  Although lithium provided only marginal added information for
most of the sample, the independent lithium and rotation ages agreed
for $\gtrsim$90\% of cases.  In particular, we reported {\bf XX}
confirmed $<$500\,Myr planets with strong age constraints from both
rotation and lithium, including {\bf YY} mini-Neptunes, {\bf ZZ}
super-Earths, {\bf AA} Earths and sub-Earths, and a Neptune-sized
planet.

Our three main conclusions with respect to our original questions are
as follows.

\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  %
  \item The age distributions of both the Kepler target stars and the
    known Kepler planet-hosts show a demographic cliff.  There are
    \ratioobtoybstars\ times as many ``old'' (2-3\,Gyr) stars in the
    Kepler field as ``young'' (0-1\,Gyr) stars.  The star formation
    rate today is \ratiosfr$\pm$\uncratiosfr\ times lower than it was
    three billion years ago.
    This star formation history broadly agrees with those derived
    from CMD fitting and white-dwarf chronology, though with
    important nuances.
  %
  \item Rather than expecting $\approx$500 exoplanets younger than one
    billion years, the age distribution of stars in the
    Kepler field implies that we should instead expect $\approx$250.
  %
  %FIXME: 120 number
  \item Using rotation-based ages, we have derived ages for
    \nplyounggyrotwosigma\ Kepler planets younger than 1\,Gyr at 2$\sigma$,
    and \nplyounggyro\ planets with median ages below 1\,Gyr.
    Concatenating these planets against the existing literature yields
    $\approx$120-170 known sub-gigayear planets with well-constrained
    ages.  This lessens the original factor of ten discrepancy to
    a factor of at most two.
\end{enumerate}

\acknowledgements
This work was supported by the Heising-Simons 51~Pegasi~b Fellowship
(LGB)
and the Arthur R.~Adams SURF Fellowship (EKP).

(author contribution statements will go here)
L.G.B.~conceived the project, collected data, 
inferred ages, and wrote the manuscript.
E.K.P.~contributed to early iterations of the rotation-based age analysis.
L.A.H.~contributed to project design.
H.I. and A.W.H~contributed to acquisition, reduction, and analysis of
the HIRES data.
All authors assisted in manuscript revision.

\facilities{
  Gaia \citep{Gaia_DR3_2022},
  Kepler \citep{Borucki10},
  Keck:I (HIRES),
	2MASS \citep{Skrutskie06},
	SDSS \citep{2000AJ....120.1579Y}.
}

\software{
  astropy \citep{Astropy18},
  matplotlib \citep{matplotlib},
  numpy \citep{numpy},
  scipy \citep{scipy},
}

\clearpage

\startlongtable
\begin{deluxetable*}{llllllllrrrrrrl}
  \tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  This
  version of the table is truncated to include only the youngest
  confirmed Kepler planets, sorted by the minimum of either the
  rotation or lithium-based age.  The full machine-readable table
  contains ages and age limits for \nnonfopkoissomeageinfo\ non-false
  positive KOIs with MES$>$10.  
  A bash script to decode any given $Q_{\rm star}$ quality flag is
  {\bf
  \href{https://gist.github.com/lgbouma/20368253f1a98da1b39cf32fdda0be13}{available
  online}}.  A python script to select stars with specific bit flags
  is {\bf \href{https://gist.github.com/lgbouma/87ad8bde42625e766a5a8857cc5a183a}{also available}}.
  \label{tab:planets}}
  %\toprule
  %\midrule
  %\endhead
  \startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ &
  $t_{\rm gyro}$ & $t_{\rm Li} $ & Consistent? & $R_{\rm p}$ & $P_{\rm orb}$ &
  $Q_{\rm planet}$ & $Q_{\rm star}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &  str &   Earths &    days &       int  & int & bool & -- \\
  \hline
  \input{table_subgyr_confirmed} %\\
% \rule
%  \input{table_subgyr_candidate}
  \enddata
  \tablecomments{Two checkmarks denotes systems with two-sided $t_{\rm
  gyro}$ and $t_{\rm Li}$ for which no important quality flags are
  raised.  One checkmark denotes two-sided $t_{\rm gyro}$ only. 
  ``Confirmed'' planets appear in the machine-readable
  version before``candidate'' planets.  The bit quality flags for the
  rotation-based ages, $Q_{\rm star}$, are described in
  Section~\ref{subsec:flags}.  Concisely summarized, they are:
  Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
  Bit 1: $\log g$$<$4.2?
  Bit 2: $M_{\rm G}$<3.9 or $M_{\rm G}>8.5$?
  Bit 3: In KEBC?
  Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
  Bit 5: Confused Kep-Gaia crossmatch?
  Bit 6: Gaia DR3 non-single star?
  Bit 7: RUWE$>$1.4?
  Bit 8: Crowded?
  Bit 9: Far from main sequence?
  Bit 10: \citetalias{Santos_2021} CP/CB?
  Bit 11: $P_{\rm rot}$ not in the homogeneous
  \citetalias{Santos_2019} \citetalias{Santos_2021} sample?
  As an example, Kepler-1627 has $Q_{\rm star}$ flagged with bit 10 and bit 7.
  The analogous planet quality bitmask, $Q_{\rm planet}$, has the
  following meaning.
  Bit 0: Candidate not reliable?
  Bit 1: MES$<$10?
  Bit 2: Grazing?
  For a star to have a high likelihood of being ``reliable for
  gyrochronology'' we suggest $Q_{\rm star}$ bits zero to nine to
  not be raised, and for a planet to be ``reliabe'', we suggested
  $Q_{\rm planet}$ to be zero.
  }
\end{deluxetable*}

\startlongtable
\begin{deluxetable*}{lllllllrrrrrr}
  \tabletypesize{\scriptsize}
  \tablecaption{
  Ages of Kepler target stars derived from rotation periods.  The
  full machine-readable table includes \nuniqstarsantosrot\ stars
  with reported rotation periods from \citetalias{Santos_2019} and
  \citetalias{Santos_2021}, \nuniqstarfinitegyroage\ of which have
  finite reported gyrochrone ages.  A random set of these stars is
  shown here to give an idea for form and content.  The quality
  bitmask is as in Table~\ref{tab:planets}; requiring bits zero to
  nine to be null yields \nuniqstarsantosrotgyroappl\ stars for
  which gyrochronology is likely to be valid.  \label{tab:stars}}
  %\toprule
  %\midrule
  %\endhead
  \startdata
  KIC & Gaia DR3 &  $T_{\rm eff}$ & $P_{\rm rot}$ & $t_{\rm gyro}$ & $Q_{\rm star}$ \\
  -- &   -- & K & days &  Myr &    int  \\
  \hline
  \input{table_star_gyro} %\\
  %\hline
  \enddata
% \tablecomments{This table includes X, Y, Z... The machine-readable version,
%   available online, includes additional columns for .... 
% }
\end{deluxetable*}

%\clearpage

\bibliographystyle{aasjournal}
\bibliography{bibliography}

\appendix
\section{Cases With ``Inconsistent'' Rotation and Lithium-Based Ages}
\label{app:inconsistent}

In this section we discuss the systems with confirmed planets that
have nominally discrepant rotation and lithium-based ages.  We find
that only one of them, Kepler-786, is interesting.

\paragraph{Kepler-1}---TrES-2/Kepler-1 \citep{2006ApJ...651L..61O} has
a G0 V primary and a $\approx$2.5\,day near-grazing transiting hot
Jupiter, and has been studied in detail by multiple investigators
\citep[e.g.][]{2007ApJ...664.1190S,2008ApJ...682.1283W,2011ApJ...733...36K,2011MNRAS.417.2166S}.
Our lithium-based age of this system (\trestwotli\,Myr), derive from
EW$_{\rm Li}$=81$\pm$3\,m\AA, is qualitatively in line with the
previously noted lithium abundance \citep{2007ApJ...664.1190S}.
However, we detect no photometric rotation signal, and the
spectroscopic $v\sin i$ is low.  The lithium age is strongly
asymmetric because there is significant amounts of scatter in EW$_{\rm
Li}$ vs.~$T_{\rm eff}$ over ages $t$$\gtrsim$1\,Gyr for early G
dwarfs.  We calculate a $2\sigma$ upper limit on $t_{\rm Li}$ of
2.8\,Gyr, and a $3\sigma$ upper limit of 6.1\,Gyr.  The non-detection
of rotation, while mildly surprising, is not shocking if the system's
true age is in this long tail.

\paragraph{Kepler-101}---This $\approx$5500\,K star has EW$_{\rm
Li}$=100$\pm$5\,m\AA, nominally implying $t_{\rm
Li}$$\approx$200-800\,Myr, but no rotation detection.  Our automated
flags noted that this star has a low surface gravity, high luminosity,
and is far from the main sequence.  \citet{2014A&A...572A...2B}
concur: this star is a sub-giant slightly more massive than the Sun;
no rotation detection is expected; the lithium likely simply survived
over the star's main sequence lifetime.

\paragraph{Kepler-108}---This system of two mutually inclined giant
planets \citep{2017AJ....153...45M} has a $T_{\rm
eff}$$\approx$5600\,K host star with a strong lithium detection, and
no rotation detection.  The host star is entering the red giant
branch; this has been previously noted through asteroseismology
\citep{2013ApJ...767..127H}, and our analysis of the Keck/HIRES
spectrum with \texttt{SpecMatch-Synth} \citep{2017AJ....154..107P}
concurs, yielding $T_{\rm eff}$=5668$\pm$100\,K, $\log g$=3.8$\pm$0.1,
[Fe/H]=0.38$\pm$0.06, and $v\sin i$=3.2$\pm$1.0\kms.  Given the star's
high mass, the high lithium content and lack of rotation detection is
not surprising.

%Independently, Kepler~108 was also reported by \citet{2020ApJ...903...55P} to be a member
%of NGC6811.  The cluster has $\mu_\alpha\approx-3.4$mas/yr and
%$\mu_\delta \approx -8.8$mas/yr.  Kepler~108 has
%$\mu_\alpha=-1.5\pm0.1$\,mas/yr and $\mu_\delta \approx
%-1.2\pm0.1$\,mas/yr \citep{Gaia_DR3_2022}.  We therefore suggest that it is not a member
%of NGC6811.

\paragraph{Kepler-786}---This K3 V star has a surprisingly high
lithium content.  The spectrum is single-lined, with
\texttt{SpecMatch-Synth} derived parameters of $T_{\rm
eff}$=4769$\pm$100\,K, $\log g$=4.5$\pm$0.1, [Fe/H]=0.11$\pm$0.06, and
$v\sin i$=2.4$\pm$0.6\kms.  Yet with EW$_{\rm Li}$=83$\pm$6\,m\AA, the
lithium age of \kepseveneightsix\,Myr would predict an obvious
rotation signal.  None is present in the Kepler light curve,
consistent with the low $v\sin i$.  Out of all ``apparently
discrepant'' lithium and rotation measurements discussed in this
appendix, this is the only one that seems to remain discrepant after
scrutiny.  The  \ion{Ca}{2} doublet is in emission, with $R'_{\rm
HK}=-4.7\pm0.5$, which suggests an age at least as old as the Hyades
\citep{Mamajek_2008}.  The Balmer lines are in absorption, and display
no obvious signatures of youth.

\paragraph{Kepler-1644}---The rotation-based age of this system
(\kepsixteenfourfour\,Myr) is nominally much lower than the lithium
limit ($>$767\,Myr).  The Kepler light curve shows a $\approx$1\%
amplitude 1.4\,day rotation signal with many flares.  However the
automated quality flags note that the star has low (photometric)
surface gravity, a high RUWE (${\rm RUWE}_{\rm DR3}$=7.0), and that it
is far from the main sequence.  The Keck/HIRES spectrum also shows
visually narrow lines, with a \texttt{SpecMatch-Synth} $v \sin i \leq
2$\kms.  However, we performed a cross-correlation between the HIRES
spectrum and the nearest matches in the Keck/HIRES template library
\citep{2015AJ....149...18K}, and found that on 31 July 2022 (UT) the
system displayed a broad CCF with a blended second component at
$\approx$$+26$\,\kms relative to the primary, with a best-fit
flux-ratio of $\approx$3.4\%, and a preferred $T_{\rm
eff,B}$$\approx$4400\,K.  The spectrum and astrometric excess noise
therefore point to this system being an unresolved binary, which calls
the reliability of the rotation-based age into question.  The
non-detection of the companion from Robo-AO imaging at Palomar
\citep{2017AJ....153...66Z} suggests that the companion(s) are likely
within $\rho$$\lesssim$0.3$''$.

\paragraph{Kepler-1699}---This system is in a similar qualitative
regime as Kepler-1644, with an apparently young $t_{\rm
gyro}$=\kepsixteenninenine\,Myr derived from a $\approx$2\% amplitude
4.2\,day rotation signal, and no evidence for lithium.  The system has
${\rm RUWE}_{\rm DR3}$=19.1.  From the same style of CCF analysis from
the Keck/HIRES spectrum acquired on 31 Aug 2022 (UT), we also find a
double-peaked CCF, in this case with a secondary component at
$\approx$$-16$\,\kms relative to the primary with $T_{\rm
eff,B}$$\approx$4900\,K.  This putative companion is similarly not
detected in high-resolution imaging \citep{2017AJ....153...66Z}.  The
rotation-based age is questionable, given that we do not know which
source the rotation signal is from, or whether the stars have
interacted.

\paragraph{Kepler-1943}---This system nominally has a
\kepnineteenfourthree\,Myr lithium age, and no reported rotation
detection.  However, the star is flagged as being over-luminous,
low-surface gravity, and far from the main sequence.  In other words,
it is a subgiant.

\paragraph{Kepler-639, Kepler-320, Kepler-1719, Kepler-1876,
Kepler-1072, Kepler-1743, Kepler-1929, Kepler-1488}---These eight
systems all have nominally two-sided lithium age posteriors between 1
and 3\,Gyr, and yet lack rotation period detections.  All are
subgiants, flagged with $Q_{\rm star}$ under various combinations of
bits 1, 2, and 9.

\section{Notes on Individual Systems}

%Kepler~221 is a 633$^{+120}_{-112}$\,Myr four-planet system.
%
%Kepler~51 (CITE Masuda2014) has previously been noted for its bizarre architure.
%The rotation-based age we find here is higher/lower than previously noted.

Kepler~411, a $\approx$780$\pm$170\,Myr early K dwarf with three
transiting planets at 3.0, 7.8, and 58\,days, has a rare
architecture.  \citet{2019A&A...624A..15S} conducted a transiting
timing analysis of the system, confirming Kepler-411d and arguing for
the presence of a fourth non-transiting planet, Kepler-411e, based on
the TTVs seen in Kepler-411d and assuming that the system is nearly
coplanar.  Regarding the stellar age, \citet{2019A&A...624A..15S} used
the \citet{2007ApJ...669.1167B} relation and the measured rotation
period, which yielded an estimate of 212$\pm$31\,Myr.  However, the
star's temperature ($\approx$4900\,K) puts it in a regime where the
\citeauthor{2007ApJ...669.1167B} relation has been known to not fit
observations for a while
\citep[e.g.][Fig.~9]{Mamajek_2008}.  Our adopted age of
$\approx$780$\pm$170\,Myr agrees with the open cluster calibrators.

\section{TODO's}

\begin{enumerate}
%  \item Compare Mathur2023 ages against my ages, and (more interestingly) against the
%    cluster scale using NGC6811, Theia520, Melange3, and CepHer stars.
%    Could compare against asteroseis too.
%    %
%  \item Intro needs better discussion of Mathur2023 and Lu2021
    %
  \item Consider moving the bit list description to its own appendix.
    %
  \item Rerun gyro calculation with denser grid, fixing quantization
    and extending upper limit to 5gyr.  (Will use updated Teff
    uncertainties)
\end{enumerate}

%\section{Comparison Against Previous Literature Age Scales}
%Searching the literature for gyrochronal analyses of the Kepler field,
%the most relevant studies seemed to be those of
%\citet{Walkowicz_2013}, \citet{Reinhold_2015}, and 
%\citet{David_2021}.
%
%Also Mathur2023, which does ``magneto-gyrochronology'', including the
%Sph indicator in a model of the ages.
%
%
%\subsection{Asteroseismic Age Comparison}
%T Ceillier, J Van Saders et al 2016 MNRAS...

%\section{Figure dump}
%
%\begin{figure*}[!t]
%  \begin{center}
%    \leavevmode
%    \includegraphics[width=0.5\textwidth]{gyroage_vs_teff_errs_showplanets_linear.png}
%  \end{center}
%  \vspace{-0.6cm}
%  \caption{
%    {\bf Gyro-age vs Teff}.
%    $\pm$1$\sigma$ uncertainties are plotted around the median for Kepler stars (black) and KOIs (blue).
%    Visual pileup on the left is b/c i) uncertain ages, and ii) lots of stars there.  Precision-land is the other portion.
%    \label{fig:gyroage_vs_teff}
%  }
%\end{figure*}

\clearpage
\listofchanges

\end{document}